---
title: "Personal activity"
author: "Rhao"
date: "Thursday, May 07, 2015"
output: html_document
---

```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
# make this an external chunk that can be included in any file
library(knitr)
options(width = 100)
opts_chunk$set(message = F, error = F, warning = F, comment = NA, fig.align = 'center', 
               dpi = 100, tidy = F, cache.path = '.cache/', fig.path = 'fig/')
```


#Introduction
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 
The project consists of building a machine learning algorithm to predict activity quality from activity monitors.

#Getting data
The data is read here and training and testing set are created
```{r}
library(caret)
library(rattle)
library(randomForest)

alldata = read.csv("pml-training.csv")
set.seed(32323)

inTrain <- createDataPartition(y=alldata$classe,p=0.7, list=FALSE) 
training <- alldata[inTrain,]
testing <- alldata[-inTrain,]
```


#Cleaning up data
Based on the summary of the data only colums that have useful values and create dataset with these columns
```{r, results="hide"}
dim(training)
summary(training)
list_of_cols <- c(
  "roll_belt"          ,  "pitch_belt"          , "yaw_belt"            , "total_accel_belt",    
  "gyros_belt_x"      ,   "gyros_belt_y"        , "gyros_belt_z"        , "accel_belt_x"    ,   
  "accel_belt_y"      ,   "accel_belt_z"        , "magnet_belt_x"       , "magnet_belt_y"   ,    
  "magnet_belt_z"     ,   "roll_arm"            , "pitch_arm"           , "yaw_arm"         ,    
  "total_accel_arm"   ,   "gyros_arm_x"         , "gyros_arm_y"         , "gyros_arm_z"     ,    
  "accel_arm_x"       ,   "accel_arm_y"         , "accel_arm_z"         , "magnet_arm_x"    ,    
  "magnet_arm_y"      ,   "magnet_arm_z"        , "roll_dumbbell"       , "pitch_dumbbell"  ,    
  "yaw_dumbbell"      ,   "total_accel_dumbbell", "gyros_dumbbell_x"    , "gyros_dumbbell_y",    
  "gyros_dumbbell_z"  ,   "accel_dumbbell_x"    , "accel_dumbbell_y"    , "accel_dumbbell_z",    
  "magnet_dumbbell_x" ,   "magnet_dumbbell_y"   , "magnet_dumbbell_z"   , "roll_forearm"    ,    
  "pitch_forearm"     ,   "yaw_forearm"         , "total_accel_forearm" , "gyros_forearm_x" ,    
  "gyros_forearm_y"   ,   "gyros_forearm_z"     , "accel_forearm_x"     , "accel_forearm_y" ,   
  "accel_forearm_z"   ,   "magnet_forearm_x"    , "magnet_forearm_y"    , "magnet_forearm_z");    
testingclean <- testing[ , which(names(testing) %in% list_of_cols)] 
trainingclean <- training[ , which(names(training) %in% list_of_cols)] 

```

#Run different models
##Model 1 Tree
```{r}
modelFit1 <- train(training$classe ~ ., data=trainingclean, method="rpart")
modelFit1
fancyRpartPlot(modelFit1$finalModel)
```

##Model 2 Random Forest
```{r}
modelFit2  <-randomForest(training$classe ~ ., 
                   data=trainingclean, 
                   prox=TRUE, 
                   ntree=20, 
                   tuneGrid = data.frame(mtry = 3))

modelFit2
head(getTree(modelFit2,k=2))
varImpPlot(modelFit2)
```

#Predict testing values and getting accuracy
```{r}
confusionMatrix(testing$classe,predict(modelFit1,testingclean))
confusionMatrix(testing$classe,predict(modelFit2,testingclean))
print("Accuracy for model 1:")
res1 <- confusionMatrix(testing$classe,predict(modelFit1,testingclean))
res1$overall[1]
print("Accuracy for model 1:")
res2 <- confusionMatrix(testing$classe,predict(modelFit2,testingclean))
res2$overall[1]
```

#Conclusion 
The second model done with random forest should used as it's accuracy is very high: `r I(res2$overall[1])` It is
a great machine learning algorithm to predict activity quality from activity monitors.


